### 2016-06-10 paper.tex
paper.{tex,bib} - fix review comments

\documentclass[11pt]{article}
\usepackage{acl2016}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}

%\aclfinalcopy % Uncomment this line for the final submission
\input{bit-aclfinalcopy}
\def\aclpaperid{16} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.
\usepackage[noinline,nomargin,noindex,draft]{fixme}
\fxsetup{theme=color,mode=multiuser}
\FXRegisterAuthor{n}{note}{NOTE} 
%\nnote[inline]{this is how to make regular notes...}
%\nerror[margin]{} or \fxerror[inline]{reserved for final paper}

\newcommand\BibTeX{B{\sc ib}\TeX}

\title{bot.zen @ EmpiriST 2015 - A minimally-deep learning PoS-tagger \\ (trained for German CMC and Web data)}

\author{Egon W.~Stemle \\
  EURAC \\
  Bozen-Bolzano, Italy \\
  {\tt egon.stemle@eurac.edu}} % \\\And
  %Jennifer-Carmen Frey \\
  %EURAC \\
  %Bozen-Bolzano, Italy \\
  %{\tt jennifercarmen.frey@eurac.edu} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}
    This article describes the system that participated in the Part-of-speech
    tagging subtask of the \emph{EmpiriST 2015 shared task on automatic
    linguistic annotation of computer-mediated communication / social
    media}.

    The system combines a small assertion of trending techniques, which
    implement matured methods, from natural language processing to achieve
    competitive results on POS tagging of German CMC and Web corpus data; in
    particular, the system uses word embeddings and character-level
    representations of word beginnings and endings in a LSTM recurrent neural
    network architecture.

    \fxerror[inline]{}write something about the results.

\end{abstract}

\section{Introduction}
\label{sec:intro}

Part-of-speech (POS) tagging is an essential processing stage for virtually all
NLP applications.
Subsequent tasks, like for example parsing, named-entity recognition, event
detection, and machine translation, often utilise POS tags (directly or
indirectly), and altogether, benefit from accurate tag sequences.
However, frequent phenomena in computer-mediated communication (CMC) and Web
corpora such as emoticons, acronyms, interaction words, iteration of letters,
graphostylistics, shortenings, addressing terms, and non-standard
spellings~\cite{crystal2001,androutsopoulos2007,beisswenger2013} (gravely)
deteriorate the performance of POS-taggers~\cite{giesbrecht2009,I13-1041}.
To this end, the EmpiriST shared task (ST) encouraged the developers of NLP
applications to adapt their tools and resources for the processing of written
German CMC and Web data~\cite{empirist2016}.

Recently, State-of-the-art results on various linguistic tasks were
accomplished by architectures using neural-network based word embeddings;
Baroni~\shortcite{baroni-dinu-kruszewski:2014:P14-1} conducted a set of
experiments comparing the popular
\texttt{word2vec}~\cite{DBLP:journals/corr/abs-1301-3781,arXiv:1310.4546}
implementation for creating word embeddings to other recent distributional
methods with state-of-the-art results across various (semantic) tasks. 
Even though these results suggest that the word embeddings substantially
outperform the recent architectures on semantic similarity and analogy
detection tasks Levy~\shortcite{TACL570} conducted a comprehensive set of
experiments and comparisons that suggest that much of the improved results are
due to the system design and parameter optimizations, rather than the selected
method; they conclude that "there does not seem to be a consistent significant
advantage to one approach over the other".

We take these results to use \texttt{word2vec} word embeddings in our
architecture because they are (1) often faster to learn, and (2) for
educational purposes, i.e.~we wanted to get some experience in using them.
\fixme[margin]{Line 70: "...often faster..." is there some evidence / citation for that?}
\fixme[margin]{Line 71: "...we wanted to get some experience..." -> nobody cares}

\section{Design}
\label{sec:design}
Our overall design takes inspiration from as far back as
Benello~\shortcite{Benello1989} who used a few preceding and one following word
in a feed-forward network with back-propagation for PoS tagging;
they achieved an accuracy of $0.95$ on the Brown Corpus.

Further inspiration comes from dos Santos~\shortcite{santos2014learning} who
used a deep neural network trained with character-level joint with word
embedding representation of words for PoS tagging. On the Wall Street Journal
(WSJ) portion of the Penn Treebank they achieve accuracy scores of $0.9732$.

Our design comprises character-level 3-grams of word beginnings and ends joint
with word embedding representation of words in a long short-term memory (LSTM)
recurrent neural network (RNN)
architecture~\cite{Hochreiter:1997:LSM:1246443.1246450} for learning a mapping
of sequences of these representations to PoS tags.

\section{Implementation}
\label{sec:implementation}

We maintain the implementation in a source code repository at
\url{https://github.com/iiegn/bot_zen}.  The version tagged as {\tt 0.9}
comprises the version that was used to generate the results submitted to the
ST; 
the version tagged as {\tt 1.0} is identical at its core but comes with
explicit documentation on how to download and install external software, and
how to download and pre-process required corpora.

%\subsection{Word Embeddings}
For learning word embeddings we use \texttt{word2vec}; for building our neural
netowrk we use Keras, a minimalist, highly modular neural networks library,
written in Python and capable of running on top of either TensorFlow or
Theano~\cite{chollet2015}; in our case it runs on top of Theano, a Python
library that allows to define, optimize, and evaluate mathematical expressions
involving multi-dimensional arrays
efficiently~\cite{Theano2016}.

The input to our network consists of the concatenation of the word embeddings
and the one-hot n-gram representations of the word sequences with the length of
the sequence being the whole sentence; this input is feed into a LSTM layer
that in turn projects into a fully connected output layer with softmax
activation function.

\section{Case Studies}
\label{sec:casestudies}

The implementation from Section~\ref{sec:implementation} was used to
participate in the EmpiriST 2015 shared task (see
Subsection~\ref{ssec:case_empirist}), and subsequently, to run experiments with
different data and parameters for training (see
Subsection~\ref{ssec:case_postempirist}).

\emph{empirist}%
\footnote{\url{https://sites.google.com/site/empirist2015/home/shared-task-data}}
is the CMC and Web training data containing data samples from different CMC
genres and samples from text genres on the Web.
The data were tagged by two annotators, and unclear cases were decided by a
third person. See Bei{\ss}wenger~\shortcite{empirist2016} for more details.

\emph{Tiger v2.2}%
\footnote{\url{http://www.ims.uni-stuttgart.de/forschung/ressourcen/korpora/tiger.html}}
is version 2.2 of the TIGERCorpus~\cite{Brants2004} containing German newspaper
texts. 
The corpus was semi-automatically POS-tagged. For research and evaluation
purposes, the TIGERCorpus can be downloaded for free.

\emph{de.wiki'15}%
\footnote{\url{http://www1.ids-mannheim.de/kl/projekte/korpora/verfuegbarkeit.html#Download}}
are user talk pages (messages from users to users, often questions and advice)
and article talk pages (questions, concerns or comments related to improving a
Wikipedia article), and article pages of the German wikipedia from 2015, made
available by the \emph{Institut f\"{u}r Deutsche
Sprache}\footnote{\url{http://www.ids-mannheim.de}}.
The user talk pages contain
% 378,576,711 
$0.9*10^9$ tokens in
% 15,343,904
$15*10^6$ sentences, the article talk pages
% 446,551,142 
$0.4*10^9$ tokens in
%17,275,385 
$17*10^6$ sentences, and the article pages
% 1,126,462,560 
$1.1*10^9$
tokens in
% 46,740,611 
$46*10^6$ sentences.
% tokens: 378,576,711+446,551,142+1,126,462,560 = 1,951,590,413
% sentences: 15,343,904+17,275,385+46,740,611 = 79359900
The data is available under the CC BY-SA 3.0\footnote{Creative Commons
Attribution-ShareAlike 3.0 Unported, i.e.~the data can be copied and
redistributed, and adapted for any purpose, even commercially. See
\url{http://creativecommons.org/licenses/by-sa/3.0/} for more details.}
license.

\begin{table}[h]
\begin{center}
\begin{tabular}{l|r|r}
\hline
\multicolumn{1}{c}{}	& \multicolumn{1}{c}{Tokens}	& \multicolumn{1}{c}{Sentences} \\ \hline
empirist    			& $10*10^3$	& $0.6*10^3$      \\ \hline
Tiger v2.2 		& $0.9*10^9$	& $0.05*10^6$	  \\ \hline
de.wiki'15   	& $2.0*10^9$	& $79*10^6$	  \\ \hline\hline
C4Corpus            	& $$		& $$	  \\ \hline
DECOW14         & $11.7*10^9$	& $600*10^6$	  \\ \hline
deTenTen	    & $19.9*10^9$	& $1300*10^6$	  \\ \hline
\end{tabular}
\end{center}
\caption{\label{tab:corpora}Training data used for EmpiriST (upper part) and
additional data for the subsequent experiments (lower part).} \end{table}

\emph{C4Corpus}%
\footnote{\url{https://github.com/dkpro/dkpro-c4corpus}} is a full documents
German Web corpus that has been extracted from CommonCrawl, the largest
publicly available general Web crawl to date. 
The data is available under the CreativeCommons license
family\footnote{\url{https://creativecommons.org/licenses/}}. 
See Habernal~\shortcite{Habernal.et.al.2016.LREC} for details about the corpus
construction pipeline, and other information about the corpus.

\emph{DECOW14}%
\footnote{\url{http://corporafromtheweb.org/decow14/}} is the sentence shuffled
German Web corpus by COW created with the 2014 technology of the COW
initiative~\cite{SchaeferBildhauer2012}.
The data can be used for purposes of academic research.
% $11.7*10^9$ tokens 
% $0.6*10^9$ (624,767,747) sentences

\emph{deTenTen} Web corpus.
%DeTenTen
% $19.9*10^9$ (19,918,263,493) tokens
% $1.3*10^9$ (1,309,696,220) sentences
\cite{1120431}

An overview of the used training data can be found in Table~\ref{tab:corpora}.
\fixme[margin]{Line 187: as far as I know the trial data is part of the training data. Testing
on this will naturally yield good, you are training on your evaluation data.
What is the performance on the provided test set?}

For evaluation we used the EmpiriST 2015 trial data: 4000 tokens, PoS tagged by
one annotator without systematic error checks, i.e.~the accuracy of the PoS
annotations is likely lower than on the training data and on the gold standard.

\subsection{EmpiriST 2015}
\label{ssec:case_empirist}

For the three submitted  runs of EmpiriST 2015 we used word embeddings from the
\emph{empirist} data (run 1), the \emph{de.wiki'15} data (run 2), and both (run
3).  We trained a 500-dimensional skip-gram model on the \emph{empirist} data
that ignored all words with less than 3 occurrences within a window size of 10;
it was trained with negative sampling (value 5) and (accidentally%
\footnote{Although, technically, the two can be combined they 'work against'
each other: hierarchical softmax works better with few training epochs but for
negative sampling, few training epochs are suboptimal.}% https://groups.google.com/forum/#!topic/word2vec-toolkit/WUWad9fL0jU
) also with hierarchical softmax.
\fixme[margin]{Line 208: you should avoid terms like accidental. Is it even possible to use
negative sampling and hierarchical softmax? To my understanding these
approaches are different techniques that cannot be combined easily. By quickly
doublechecking the word2vec source code I think only HS is used if it is
activated, negative-sampling will be disregarded in this case. The authors
should check that thoroughly though.}
We trained a 500-dimensional continuous bag-of-word model on the
\emph{de.wiki'15} data that ignored all words with less than 25 occurrences
within a window size of 10; it was trained with negative sampling (value 3) and
(accidentally) also with hierarchical softmax.

The rational behind training the two models differently was that according to
\texttt{word2vec} author's
experience\footnote{\url{https://groups.google.com/d/msg/word2vec-toolkit/NLvYXU99cAM/E5ld8LcDxlAJ}}
a skip-gram model "works well with small amount[s] of the training data, [and]
represents well even rare words or phrases", and a cbow model is "several times
faster to train than the skip-gram, [and has] slightly better accuracy for the
frequent words".
% empirist.bin -cbow 0 -size 500 -window 10 -hs 1 -negative 5 -min-count 3 
% bigdata.bin -cbow 1 -size 500 -window 10 -hs 1 -negative 3 -min-count 25
The other parameters of \texttt{word2vec} were left at its default settings.
\fixme[margin]{Line 223: what are the default settings, i.e. they might change over time, it
is better to explicitly mention them}

To optimize the output results of the neural network we ran a simple grid
search for the following three parameters: the output dimensions of the hidden
LSTM layer, the dropout value for the projections from the LSTM to the output
layer during training, and the number of epochs during training. The resulting
parameters were $1024$ output dimensions, $0.1$ as the dropout value, and $20$
number of epochs.
%lstm_output_dims = [128, 256, 512, 1024*]
%dropouts = [0.1*, 0.25, 0.5, 0.75]
%nb_epochs = [5, 10, 20*]

For the first run we trained the network on \emph{empirist}, for the second run
on \emph{Tiger v2.2}, and for the third on the concatenation of the two. % 
% 1: empirist_training (tagset:ibk)
% 2: tiger-2.2 (tagset:1999)
% 3: empirist_training + tiger-2.2 (tagset:ibk)
Our evaluation on the EmpiriST trial data resulted in these accuracy scores:
1:~$0.88$, 
2:~$0.78$, 
3:~$0.93$;
and the overall ST evaluation was:
1:~$0.84$,
2:~$0.82$,
3:~$0.88$.
\fixme[inline]{more verbose results.}

\subsection{Post-EmpiriST Experiments}
\label{ssec:case_postempirist}
As a post-EmpiriST addition to the results we wanted to (1)~remedy some errors
and oversights in our initial pre-processing and training, (2)~optimize the
learning parameters of the word embedding models and the neural network, and
(3)~train word embedding models on other data.
\fxnote[margin]{add (1)}

\fixme[margin]{When explaining numbers to the power of 10 write $9x10^8$ instead of $0.9*10^9$, also,
in running text it is advised to use "900 million" instead.}
The optimal learning parameters for the word embedding models were selected
according to a comprehensive set of experiments and comparisons by
Levy~\shortcite{TACL570}, and observations and suggestions by
Mikolov~\shortcite{arXiv:1310.4546}. 
For the post-EmpiriST experiments we used the skip-gram with negative-sampling
training method (SGNS), with a negative-sampling value of $15$ (and without
hierarchical softmax, cf.~Subsection~\ref{ssec:case_empirist}), ignored all
words with less than $100$ occurrences within a window size of $10$, and a
sub-sampling value of $10^{-5}$. 
% cbow=0, negative=15, hs=0, window=10, sample=10^-5 
% arXiv:1310.4546: -negative 15 -sample 0.00001
% TACL570: alpha=0.75 and dyn=1-> fixed in word2vec algorithm;
% There are quite a few differences between the skip-gram and the CBOW models. However, if you have a lot of training data, their performance should be comparable.\footnote{\url{https://groups.google.com/d/msg/word2vec-toolkit/NLvYXU99cAM/E5ld8LcDxlAJ}}

As additional data to train word embedding models we obtained and processed
\emph{C4Corpus}, \emph{DECOW14}, and \emph{deTenTen}.

\fxnote[inline]{Post EmpiriST experiment results / discussion and conclusion is still missing.}

\section{Conclusion \& Outlook}
\label{sec:conclusion}

\fxnote[inline]{}Pending.

%\section*{Acknowledgments}
%The acknowledgments should go immediately before the references.  Do
%not number the acknowledgments section. Do not include this section
%when submitting your paper for review.

\clearpage
% include your own bib file like this:
%\bibliographystyle{acl}
%\bibliography{acl2016}
%\nocite{*}
\bibliography{paper}
\bibliographystyle{acl2016}

%\appendix
%\section{Supplemental Material}
%\label{sec:supplemental}
%Supplement

\end{document}